---
title: "Mass measurement"
output: html_document
---


## Comparison of results with weighted mean or without

As it is described in an article about data processing, the data within each replicate of the experiment is aggregated using weighted mean. Below we present the effect it has of the results.
The whole workflow can be found in documentation, here the focus is on the first aspect of the processing.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyr)
library(dplyr)

calculate_mass <- function(dat){
  
  proton_mass <- 1.00727647
  
  dat %>%
    mutate(exp_mass = Center*z - z*proton_mass) %>%
    select(-Center, -z, -Modification, -Fragment) %>%
    group_by(Sequence, Start, End, MHP, MaxUptake, State, Exposure, Protein, File) %>%
    summarize(avg_exp_mass = weighted.mean(exp_mass, Inten, na.rm = TRUE)) %>%
    ungroup(.) %>% 
    group_by(Sequence, Start, End, MHP, MaxUptake, State, Exposure, Protein) %>%
    summarize(mass = mean(avg_exp_mass, na.rm = TRUE),
              err_mass = coalesce(sd(avg_exp_mass, na.rm = TRUE)/sqrt(sum(!is.na(avg_exp_mass))), 0),
              num = (sum(!is.na(avg_exp_mass)))) %>%
    ungroup(.) %>%
    arrange(Start, End, Start, Exposure) %>%
    as.data.frame()
  
}

calculate_mass_no_inten <- function(dat){
  
  proton_mass <- 1.00727647
  
  dat %>%
    mutate(exp_mass = Center*z - z*proton_mass) %>%
    select(-Center, -z, -Modification, -Fragment) %>%
    group_by(Sequence, Start, End, MHP, MaxUptake, State, Exposure, Protein, File) %>%
    summarize(avg_exp_mass = mean(exp_mass, na.rm = TRUE)) %>%
    ungroup(.) %>% 
    group_by(Sequence, Start, End, MHP, MaxUptake, State, Exposure, Protein) %>%
    summarize(mass = mean(avg_exp_mass, na.rm = TRUE),
              err_mass = coalesce(sd(avg_exp_mass, na.rm = TRUE)/sqrt(sum(!is.na(avg_exp_mass))), 0)) %>%
    ungroup(.) %>%
    arrange(Start, End, Start, Exposure) %>%
    as.data.frame()
  
}

```

For the analysis we use the example file from HaDeX.

```{r}
library(HaDeX)
dat <- read_hdx(system.file(package = "HaDeX", "HaDeX/data/KD_180110_CD160_HVEM.csv"))
```

As described, the first step is to transform the Center value (geometric centroid of the isotopic envelope for given peptide in given state in given timepoint), and than to aggregate the values measured for different charge values. This is all within each replicate.

The aggregation in our workflow is a weighted mean, with Inten (intensity) values as weights, as shown below.

```{r eval=FALSE}
avg_exp_mass = weighted.mean(exp_mass, Inten, na.rm = TRUE)
```

Later, the results from replicates are aggregated using mean, and their uncertainty is calculated as standard deviation of the mean. 

How the weighted mean changes the result in comparison with simple mean? Let's see.

```{r include=FALSE}
dat_no_weight <- calculate_mass_no_inten(dat) %>%
  arrange(Start, End, State, Exposure) %>%
  mutate(source = "no_weight") %>%
  select(Sequence, Start, End, State, Exposure, mass, err_mass, source)

dat_weight <- calculate_mass(dat) %>%
  arrange(Start, End, State, Exposure) %>%
  mutate(source = "weight") %>%
  select(Sequence, Start, End, State, Exposure, mass, err_mass, source)

tmp <- bind_rows(dat_no_weight, dat_weight) %>%
  gather(type, value, -Sequence, -Start, -End, -source, -State, -Exposure) %>%
  spread(source, value) %>%
  mutate(diff = (weight-no_weight))
```

Below are calculated values of mass in two approaches, and difference between them for example peptide. 

```{r}
tmp %>%
  filter(type == "mass") %>%
  select(-type) %>%
  filter(Sequence == "ARSQKSGIRLQGHF")
```

Below are calculated values of uncertainty of mass in two approaches, and difference between them for example peptide. 

```{r}
tmp %>%
  filter(type == "err_mass") %>%
  select(-type) %>%
  filter(Sequence == "ARSQKSGIRLQGHF")
```

And for the whole set of peptides, the average difference in mass:

```{r echo=FALSE}
mean(tmp[ tmp[["type"]] == "mass" , "diff"])
```
and in uncertainty:

```{r echo=FALSE}
mean(tmp[ tmp[["type"]] == "err_mass" , "diff"])
```
Below are the histograms of this differences.

```{r message=FALSE, warning=FALSE, echo=FALSE}
library(gridExtra)
library(ggplot2)

p1 <- ggplot(filter(tmp, type == "mass"), aes(diff)) +
  geom_histogram() +
  labs(title = "Differences between mass",
       x = "Difference")

p2 <- ggplot(filter(tmp, type == "err_mass"), aes(diff)) +
  geom_histogram() +
  labs(title = "Differences between uncertainties of mass",
       x = "Difference")

grid.arrange(p1, p2, ncol=2)
```

